The Role of Context in Spatial Region Identification

### 2013-11-09 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk: correct small typos

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing a cognitive systems with this ability.
\end{abstract}

\section{Introduction} 
Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.  Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.  However, many spatial terms are defined not only by their geometry, but also their context.  That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration, functional roles, and goals.  We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).

We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.  To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

We developed the study described below in order to examine the following questions about CDSRs:
\begin{enumerate}
	\item Given a particular context, do people agree on the location and extent of a named CDSR?
	\item Does changing the context in a given scenario predictably effect the CDSR?
\end{enumerate}

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g. does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts. Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stating there was no such region).
	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).

\begin{figure}
%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
\centerline{
\begin{tabular}{cccc}
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle.pdf}\caption{}\label{fig:classroom-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-photo.jpg}\caption{}\label{fig:classroom-b}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-shaped.pdf}\caption{}\label{fig:classroom-c}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-photo.jpg}\caption{}\label{fig:classroom-d}\end{subfigure}\\
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-close.pdf}\caption{}\label{fig:classroom-f}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-far-away.pdf}\caption{}\label{fig:classroom-e}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-1.jpg}\caption{}\label{fig:bottleneck-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2.jpg}\caption{}\label{fig:bottleneck-b}\end{subfigure}\\
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close.jpg}\caption{}\label{fig:tanks-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-far.jpg}\caption{}\label{fig:tanks-b}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets.jpeg}\caption{}\label{fig:turrets-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets-plus-one.jpeg}\caption{}\label{fig:turrets-b}\end{subfigure}
\end{tabular}
}
\caption{Experiment Stimuli}
\label{fig:exp_stimuli}
\end{figure}

%\subsection{Participants and Procedure}
%Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions. 43 undergraduate students from two different third level institutes took part in the study. To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.

\textbf{Participants and Procedure:} The experiment was delivered online. 43 undergraduate students from two third level institutes took part in the study. To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.

\subsection{Task 1 - Design}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a "no region" option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.

\subsection{Task 1 - Results}
We use two methods to analyze the results of the region drawing task.  First, we use interannotator agreement for two reasons: (1) fair to good overall agreement, would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.

@@

We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation, if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli. 
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
%\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}

\begin{table}[t]
\parbox{.50\linewidth}{
\centering 
\begin{tabular}{l c c}
% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
 & Average & No \\
Stim ID & Pairwise $\kappa$ & Reg.\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
\label{tab:stimuli-kappa-scores}
}
\hfill
\parbox{.50\linewidth}{
\centering
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
maxX(f) > maxX(e) & 176.1 > 167.1 \\
% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
minX(e-flipped)  &  \\
 > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
maxY(j) > maxY(k) & 265.04 > 251.71 \\
maxY(k) = maxY(l) & 265.04 != 290.03 \\
\hline
\end{tabular}
\caption{Comparison of the expected relationships to the mean human responses.}
\label{tab:region-location-tests}
}
\end{table}

\footnotetext[2]{Significance value from T-Test (p<0.1)}

%\begin{table}
%\centerline{
%\begin{scriptsize}
%\begin{tabular}{l c c c c c c c c c c c c c}
%% (h) - 1_2; (g) - 1_1; (k) - 3_3; (f) - 2_4; (i) - 3_1; (e-flipped) - 2_3; (j) - 3_2; (f-flipped) - 2_5; (e) - 2_2;  (l) - 3_4;  (c) -  2_6;  (c-flipped) - 2_7;  (a) - 2_1
%Stim ID & (h) & (g) & (k) & (f) & (i) & (e-flipped) & (j) & (f-flipped) & (e) & (l) & (c) & (c-flipped) & (a)\\ 
%Average Pairwise $\kappa$ & 0.5034 & 0.4585 & 0.4548 & 0.3930 & 0.3906 & 0.3813 & 0.3787 & 0.3720 & 0.3629 & 0.3164 & 0.2438 & 0.2414 & 0.0996 \\ 
%No Region Responses & 0 & 1 & 0& 0& 0& 2&1&2& 1& 0&2&4& 10\\
%\hline
%\end{tabular}
%\end{scriptsize}
%}
%\label{tab:stimuli-kappa-scores}
%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimuli. The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
%\end{table}

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35. Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.

@@

%\begin{figure}
%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
%\centerline{
%\begin{tabular}{ccc}
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
%\end{tabular}
%}
%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
%\label{fig:type-1-hists}
%\end{figure}

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.  Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).  That is, a single turret does not provide safety from three tanks.  A possible explanation for this is that there is a domain expertise component to understanding contexts.

@@

\subsection{Task 2 - Design} 
In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it lay behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.

@@

%In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region. In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.

>>>>

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing cognitive systems with this ability.
\end{abstract}

\section{Introduction} 
Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.  Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.  However, many spatial terms are defined not only by their geometry, but also by their context.  That is, to understand this language requires considering the other objects and agents in the environment as well as their configuration, functional roles, and goals.  We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).

We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

Consider the following regions shown in Figure \ref{fig:examples}: the front of a classroom, safety in a military engagement, and a geographic bottleneck.  To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

We developed the study described below in order to examine the following questions about CDSRs:
\begin{enumerate}
	\item Given a particular context, do people agree on the location and extent of a named CDSR?
	\item Does changing the context in a given scenario predictably affect the CDSR?
\end{enumerate}

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given environment (e.g., does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts. Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follows the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimulus that defined the CDSR (or stating there was no such region).
	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point Likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:turrets-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).

\begin{figure}
%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
\centerline{
\begin{tabular}{cccc}
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle.pdf}\caption{}\label{fig:classroom-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-photo.jpg}\caption{}\label{fig:classroom-b}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-shaped.pdf}\caption{}\label{fig:classroom-c}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-photo.jpg}\caption{}\label{fig:classroom-d}\end{subfigure}\\
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-close.pdf}\caption{}\label{fig:classroom-f}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-far-away.pdf}\caption{}\label{fig:classroom-e}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-1.jpg}\caption{}\label{fig:bottleneck-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2.jpg}\caption{}\label{fig:bottleneck-b}\end{subfigure}\\
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close.jpg}\caption{}\label{fig:tanks-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-far.jpg}\caption{}\label{fig:tanks-b}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets.jpeg}\caption{}\label{fig:turrets-a}\end{subfigure}&
\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets-plus-one.jpeg}\caption{}\label{fig:turrets-b}\end{subfigure}
\end{tabular}
}
\caption{The stimuli used in the experiment.}
\label{fig:exp_stimuli}
\end{figure}

%\subsection{Participants and Procedure}
%Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions. 43 undergraduate students from two different third-level institutes took part in the study. To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimulus presented within each trial.

\textbf{Participants and Procedure:} The experiment was delivered online. 43 undergraduate students from two third-level institutes took part in the study. To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.

\subsection{Task 1 - Design}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.

\subsection{Task 1 - Results}
We use two methods to analyze the results of the region drawing task.  First, we use interannotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.

@@

We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimulus basis. In this calculation, if either of the subjects in a pair entered a no region response for a particular stimulus a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimulus, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimulus. 
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
%\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}

\begin{table}[t]
\parbox{.50\linewidth}{
\centering 
\begin{tabular}{l c c}
% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
 & Average & No \\
Stim ID & Pairwise $\kappa$ & Reg.\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
\label{tab:stimuli-kappa-scores}
}
\hfill
\parbox{.50\linewidth}{
\centering
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
maxX(f) > maxX(e) & 176.1 > 167.1 \\
% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
minX(e-flipped)  &  \\
 > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
maxY(j) > maxY(k) & 265.04 > 251.71 \\
maxY(k) = maxY(l) & 265.04 != 290.03 \\
\hline
\end{tabular}
\caption{Comparison of the expected relationships to the mean human responses.}
\label{tab:region-location-tests}
}
\end{table}

\footnotetext[2]{Significance value from t-test (p<0.1)}

%\begin{table}
%\centerline{
%\begin{scriptsize}
%\begin{tabular}{l c c c c c c c c c c c c c}
%% (h) - 1_2; (g) - 1_1; (k) - 3_3; (f) - 2_4; (i) - 3_1; (e-flipped) - 2_3; (j) - 3_2; (f-flipped) - 2_5; (e) - 2_2;  (l) - 3_4;  (c) -  2_6;  (c-flipped) - 2_7;  (a) - 2_1
%Stim ID & (h) & (g) & (k) & (f) & (i) & (e-flipped) & (j) & (f-flipped) & (e) & (l) & (c) & (c-flipped) & (a)\\ 
%Average Pairwise $\kappa$ & 0.5034 & 0.4585 & 0.4548 & 0.3930 & 0.3906 & 0.3813 & 0.3787 & 0.3720 & 0.3629 & 0.3164 & 0.2438 & 0.2414 & 0.0996 \\ 
%No Region Responses & 0 & 1 & 0& 0& 0& 2&1&2& 1& 0&2&4& 10\\
%\hline
%\end{tabular}
%\end{scriptsize}
%}
%\label{tab:stimuli-kappa-scores}
%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimulus. The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
%\end{table}

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimulus used in the experiment and the number of no region responses for each stimulus.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flipped), and the circle layout have kappa scores of less than 0.35. Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is affected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.

@@

%\begin{figure}
%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
%\centerline{
%\begin{tabular}{ccc}
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
%\end{tabular}
%}
%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimulus, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
%\label{fig:type-1-hists}
%\end{figure}

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.  Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected relationship was the safety region between the row of turrets (k) and turrets +1(l).  That is, a single turret does not provide safety from three tanks.  A possible explanation for this is that there is a domain expertise component to understanding contexts.

@@

\subsection{Task 2 - Design} 
In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.  In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j) and (k) the point lay between all forces, in stimulus (i), it lay behind the friendly tanks, and in stimulus (l) , it lay behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.

@@

%In the type 1 trials subjects drew a polygon on the visual stimulus that defined the location and extent of the relevant spatial region. In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.

%The type 2 trials the same visual stimulus and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale their judgement of how well the location matched %the described spatial region. For each stimulus domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
